Toronto's Reaction to Fight Against Global Warming

### 2022-04-23 outputs/paper/paper.Rmd
finished everything other than discussion

```{r setup, include = FALSE}
# Import all necessary libraries
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(plotrix)
# Read the dataset, use here to navigate to working directory
global_avg_temperature <- read_csv(here::here("inputs/data/GlobalTemperatures.csv"))

#annual_energy_2011 <- read_csv(here::here("inputs/data/energy-2011.csv"))
#annual_energy_2012 <- read_csv(here::here("inputs/data/energy-2012.csv"))
#annual_energy_2013 <- read_csv(here::here("inputs/data/energy-2013.csv"))
#annual_energy_2014 <- read_csv(here::here("inputs/data/energy-2014.csv"))
#annual_energy_2015 <- read_csv(here::here("inputs/data/energy-2015.csv"))
#annual_energy_2016 <- read_csv(here::here("inputs/data/energy-2016.csv"))
#annual_energy_2017 <- read_csv(here::here("inputs/data/energy-2017.csv"))
#annual_energy_2018 <- read_csv(here::here("inputs/data/energy-2018.csv"))
climate_perception <- read_csv(here::here("inputs/data/survey.csv"))
```

@@

## Data Source and Collection
The dataset I will be using are from the City of Toronto Data Portal, it is conducted city-wide to gather people's opinion on climate change [@opendatator], the goal of this survey is to capture their perceptions about climate change and to estimate people's willingness to take actions and help the city government to establish the goal of achieving net-zero carbon dioxide emission. The survey is conducted online and a total of 404 results were received between October 11 to October 18, 2018. All survey respondents are adult residents living in the City of Toronto. All respondents are from 4 geological areas: Etobicoke-York, North York, Scarborough, and Toronto and East York, Figure \@ref(fig:respondents-number) is a bar graph showing the number of respondents from each area.

```{r respondents-number, fig.cap='Number of Respondents from Each Area', echo=FALSE, warning=FALSE, message = FALSE, fig.width=5, fig.height=3, fig.pos = "!H"}
number_from_each_area <- c(99, 101, 91, 113)
area_name <- c("Etobicoke-York", "North York", "Scarborough", "Toronto and East York")
ggplot() + geom_bar(aes(x=area_name,y=number_from_each_area), stat='identity') + xlab('Area') + ylab('Number of Respondents')
```

Other than coming from different city areas, respondents also have different demographic backgrounds, For instance, some respondents have a household income of more than $150000 while some respondents only have a household income lower than $40000. The respondents also have a big age variety with some people being 65 years old and some only 20 years old. Below are some figures showing the distribution of the respondents based on their demographic factors. Figure \@ref(fig:agesum) shows the distributions of respondents based on their age, Figure \@ref(fig:incomesum) shows the distribution of respondents based on their household income. Finally, Figure \@ref(fig:edulevel) shows the distribution of respondents based on their level of education.

@@

```{r incomesum, fig.cap='Number of Respondents Based on Household Income', echo=FALSE, warning=FALSE, message = FALSE, fig.width=5, fig.height=3, fig.pos = "!H"}

@@

@@

Our ultimate goal is to develop a statistical model which takes a person's demographic factor as input, and try to predict whether that person will be concerned about global climate change. To do this, we will first take a look at different demographic factors, and determine whether it is relevant.

@@

We will now take a look at our survey results and determine whether age is a relevant factor. The second question in the survey is "For each of the following, how concerned are you about climate change?", the respondents can choose one from the following: "Extremely concerned, Very concerned, Not very concerned, Not at all concerned, Don’t know". For simplicity, we will group *Extremely concerned* and *Very concerned* into a big category called *Concerned*, and we will group *Not very concerned* and *Not at all concerned* into a big category called *Not Concerned*, we will neglect *Don't know* answers. Figure \@ref(fig:ageconcern) is a summary table.

@@

However, the survey suggests something different, we can see that people who is older than 65 years old are much more concerned about climate change than the younger generations.

@@

We will also take a look at how household income affect people's opinion on global climate change. In this survey, income are separated into 6 different categories: *Under $40,000*, *$40,001 to $60,000*, *$60,001 to $80,000*, *$80,001 to $100,000*, *$100,001 to $150,000*, *More than $150,000*. Because this will be used as an input to the statistical model, we will simplify those ranges by taking its median, therefore, those range will become: *$30000*, *$50000*, *$70000*, *$90000*, *$125000*, *$200000*. (The Under 50000 and over 150000 are slightly adjusted). Figure \@ref(fig:incomeconcern) is a summary table.

@@

From the table, residents that are most concerned about climate change are people with household income between 80K and 100K.

# Discussion

>>>>

```{r setup, include = FALSE}
# Import all necessary libraries
knitr::opts_chunk$set(echo = TRUE)
library(modelsummary)
library(tidyverse)
library(palmerpenguins)
library(kableExtra)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(readxl)
library(plotrix)
library(gt)
# Read the dataset, use here to navigate to working directory
global_avg_temperature <- read_csv(here::here("inputs/data/GlobalTemperatures.csv"))

climate_perception <- read_csv(here::here("inputs/data/survey.csv"))
```

@@

## Data Source and Collection
The dataset I will be using are from the City of Toronto Data Portal, it is conducted city-wide to gather people's opinion on climate change [@opendatator], the goal of this survey is to capture their perceptions about climate change and to estimate people's willingness to take actions and help the city government to establish the goal of achieving net-zero carbon dioxide emission. The survey is conducted online and a total of 404 results were received between October 11 to October 18, 2018. All survey respondents are adult residents living in the City of Toronto.

## Data Overview
All respondents are from 4 geological areas: Etobicoke-York, North York, Scarborough, and Toronto and East York, Figure \@ref(fig:respondents-number) is a bar graph showing the number of respondents from each area. Other than coming from different city areas, respondents also have different demographic backgrounds, For instance, some respondents have a household income of more than $150000 while some respondents only have a household income lower than $40000. The respondents also have a big age variety with some people being 65 years old and some only 20 years old. Below are some figures showing the distribution of the respondents based on their demographic factors. Figure \@ref(fig:agesum) shows the distributions of respondents based on their age, Figure \@ref(fig:incomesum) shows the distribution of respondents based on their household income. Finally, Figure \@ref(fig:edulevel) shows the distribution of respondents based on their level of education.

```{r respondents-number, fig.cap='Number of Respondents from Each Area', echo=FALSE, warning=FALSE, message = FALSE, fig.width=5, fig.height=2.5, fig.pos = "!H"}
number_from_each_area <- c(99, 101, 91, 113)
area_name <- c("Etobicoke-York", "North York", "Scarborough", "Toronto and East York")
ggplot() + geom_bar(aes(x=area_name,y=number_from_each_area), stat='identity') + xlab('Area') + ylab('Number of Respondents')
```

@@

```{r incomesum, fig.cap='Number of Respondents Based on Household Income', echo=FALSE, warning=FALSE, message = FALSE, fig.width=4, fig.height=4}

@@

\newpage
# Models

Since our goal is to find out the impacts of demographic factors such as age, gender, income and education on one's awareness of global climate change. We will build a statistical model which takes those demographic factor as input, and output a binary variable indicating if this person will be concerned about climate change. The model will behave like the following where $\alpha_i$ denotes the $i^{th}$ demographic factor.

$$f(\alpha_0, \alpha_1, \alpha_2, \alpha_3 ...) = 0\;or\;1$$

## Logistic Regresssion

Logistic regression is used when the dependent variable (target variable) is categorical, this fits perfectly here because our target variable is either *Concerned about Climate Change* or *Not Concerned about Climate Change*. Logistic regression is very similar to the traditional linear regression but it uses the regression formula inside the logistic function of ${\frac{e^x} {(1 + e^x)}}$, so it has the following form.

$$P = {\frac{e^{(\beta_0 + \beta_1X_1)}}{1 + e^{(\beta_0 + \beta_1X_1)}}}$$
However, in our case, since we have several independent variables, we can instead use Multiple Logistic Regression, it has the following form

$$P = {\frac{e^{(\beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + ... + \beta_iX_i)}}{1 + e^{(\beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + ... + \beta_iX_i)}}}$$

## Independent Variables

For our statistical model, we will try to predict whether a person will be concerned with climate change using 4 demographic factors: Age, Gender, Income, and Level of Education.

We will investigate on whether a person's age has any effects on shaping his/her view on the global climate change.

### Gender

Because the vast majority respondents' gender is male or female. We will neglect other gender in this paper. Because gender is used as an input to the logistic regression model, we will use 1 to represent Male and 2 to represent Female.

### Income

The survey divides people's household income into 4 subgroups: *Under $40,000*, *$40,001 to $60,000*, *$60,001 to $80,000*, *$80,001 to $100,000*, *$100,001 to $150,000*, *More than $150,000*. We will use *$30000*, *$50000*, *$70000*, *$90000*, *$125000*, *$200000* to represent them respectively.\

### Level of Education

We will divide people's level of education into 3 groups: *High School or Less*, *Undergraduate Degree*, *Post Graduate Degree/Professional School*. We will use number 1 to 3 to represent each one of them respectively.

Above is a summary of the model and indepedent variables we will be using, more details will be presented in the Results section.

We will first talk about each independent variable in more detail.

@@

Our ultimate goal is to develop a statistical model which takes a person's demographic factor as input, and try to predict whether that person will be concerned about global climate change. To do this, we will take a look at different demographic factors, and determine their impacts in the model.

@@

We will now take a look at our survey results and determine whether age is a relevant factor. The second question in the survey is "For each of the following, how concerned are you about climate change?", the respondents can choose one from the following: "Extremely concerned, Very concerned, Not very concerned, Not at all concerned, Don’t know". For simplicity, we will group *Extremely concerned* and *Very concerned* into a big category called *Concerned*, and we will group *Not very concerned* and *Not at all concerned* into a big category called *Not Concerned*, we will neglect *Don't know* answers. Table \@ref(tab:ageconcern) is a summary table. However, the table suggests something different, we can see that people who is older than 65 years old are much more concerned about climate change than the younger generations.

@@

@@

We will also take a look at how household income affect people's opinion on global climate change. In this survey, income are separated into 6 different categories: *Under $40,000*, *$40,001 to $60,000*, *$60,001 to $80,000*, *$80,001 to $100,000*, *$100,001 to $150,000*, *More than $150,000*. Because this will be used as an input to the statistical model, we will simplify those ranges by taking its median, therefore, those range will become: *$30000*, *$50000*, *$70000*, *$90000*, *$125000*, *$200000*. (The Under 50000 and over 150000 are slightly adjusted). Table \@ref(tab:incomeconcern) is a summary table. From the table, residents that are most concerned about climate change are people with household income between 80K and 100K.

@@

### Gender

We will now take a look at how people's gender affect their opinion on global climate change. The survey provides 5 options: *Woman*, *Man*, *Transgender*, *Prefer not to say*, *Not listed above*, since the vast majority respondents indicated their gender to be either *Man* or *Woman*, we will neglect the rest. For simplicity, we will use 1 to represent Male and 2 to represent Female. Table \@ref(tab:genderconcern) is a summary table. From the results, we can see that female are usually more concerned about climate change than males.

```{r genderconcern, fig.cap='Concern about Climate Change Based on Gender', echo=FALSE, warning=FALSE, message = FALSE, fig.width=5, fig.height=4, fig.pos = "!H"}

gender_relevant_data <- subset(climate_perception, select = c(`QS2`, `Q1r1`))
names(gender_relevant_data)[names(gender_relevant_data) == 'QS2'] <- 'Gender'
names(gender_relevant_data)[names(gender_relevant_data) == 'Q1r1'] <- 'Concern'

gender_relevant_data <- gender_relevant_data %>% filter(!str_detect(`Concern`, "D"))
gender_relevant_data <- gender_relevant_data %>% filter(!str_detect(`Gender`, "P"))
gender_relevant_data <- gender_relevant_data %>% filter(!str_detect(`Gender`, "T"))

gender_relevant_data$`Concern`[gender_relevant_data$`Concern`=="Extremely concerned"] <- "Concerned"
gender_relevant_data$`Concern`[gender_relevant_data$`Concern`=="Very concerned"] <- "Concerned"

gender_relevant_data$`Concern`[gender_relevant_data$`Concern`=="Not very concerned"] <- "Not Concerned"
gender_relevant_data$`Concern`[gender_relevant_data$`Concern`=="Not at all concerned"] <- "Not Concerned"

gender_relevant_data$`Gender`[gender_relevant_data$`Gender` == "Man"] <- 1
gender_relevant_data$`Gender`[gender_relevant_data$`Gender` == "Woman"] <- 2

final_gender_relevant_data <- gender_relevant_data %>% group_by(`Gender`, `Concern`) %>% 
    summarize(n=n())

final_gender_relevant_data <- group_by(final_gender_relevant_data, `Gender`) %>% transmute(Concern, percent = round(n/sum(n),2))

final_gender_relevant_data %>%
  knitr::kable(caption = "Concern about Climate Change Based on Gender",
               col.names = c("Gender", "Cocern Level", "Percentage WRT Gender"),
               align = c('l', 'l', 'l'),
               booktabs = T) %>%
  kable_styling(full_width = T, latex_options = "hold_position")
```

### Level of Education

Finally, we will look at the effect of education on people's awareness of climate change. There are 6 valid responses, for simplicity, we categorize them into 3 groups: *High School or Less*, *Undergraduate Degree*, *Post Graduate Degree/Professional School*, we will label them from 1-3 respectively. Table \@ref(tab:educoncern) is a summary table.
From the table, clearly people with a higher level of education are usually more concerned with the climate change.

```{r educoncern, fig.cap='Concern about Climate Change Based on Level of Education', echo=FALSE, warning=FALSE, message = FALSE, fig.width=5, fig.height=4, fig.pos = "!H"}
edu_relevant_data <- subset(climate_perception, select = c(`QD5`, `Q1r1`))
names(edu_relevant_data)[names(edu_relevant_data) == 'QD5'] <- 'Education'
names(edu_relevant_data)[names(edu_relevant_data) == 'Q1r1'] <- 'Concern'

edu_relevant_data <- edu_relevant_data %>% filter(!str_detect(`Education`, "Prefer not to answer"))
edu_relevant_data <- edu_relevant_data %>% filter(!str_detect(`Concern`, "D"))

edu_relevant_data$`Concern`[edu_relevant_data$`Concern`=="Extremely concerned"] <- "Concerned"
edu_relevant_data$`Concern`[edu_relevant_data$`Concern`=="Very concerned"] <- "Concerned"

edu_relevant_data$`Concern`[edu_relevant_data$`Concern`=="Not very concerned"] <- "Not Concerned"
edu_relevant_data$`Concern`[edu_relevant_data$`Concern`=="Not at all concerned"] <- "Not Concerned"

edu_relevant_data$`Education`[edu_relevant_data$`Education`=="High school or less"] <- 1
edu_relevant_data$`Education`[edu_relevant_data$`Education`=="Some community college, vocational, trade school"] <- 2
edu_relevant_data$`Education`[edu_relevant_data$`Education`=="Some university"] <- 2
edu_relevant_data$`Education`[edu_relevant_data$`Education`=="Completed community college, vocational, trade school"] <- 2
edu_relevant_data$`Education`[edu_relevant_data$`Education`=="Completed undergraduate degree"] <- 2
edu_relevant_data$`Education`[edu_relevant_data$`Education`=="Post graduate/professional school"] <- 3

final_edu_relevant_data <- edu_relevant_data %>% group_by(`Education`, `Concern`) %>% 
    summarize(n=n())

final_edu_relevant_data <- group_by(final_edu_relevant_data, `Education`) %>% transmute(Concern, percent = round(n/sum(n),2))

final_edu_relevant_data %>%
  knitr::kable(caption = "Concern about Climate Change Based on Level of Education",
               col.names = c("Education", "Cocern Level", "Percentage WRT Education"),
               align = c('l', 'l', 'l'),
               booktabs = T) %>%
  kable_styling(full_width = T, latex_options = "hold_position")
```

## Model Results

We will now fit the data, Figure \@ref(fig:figmodel) shows the table of model results, it is produced by model summary [@modelsum].

```{r, echo = FALSE, message = FALSE, warning = FALSE}

# We are using 1, 2, 3, 4 to represent categories instead of using $30000 for example.
# This is because we want all indepdent variable to be the same scale.
all_relevant_data <- subset(climate_perception, select = c(`HIDAGE1`, `QD6`, `QS2`, `QD5`, `Q1r1`))
names(all_relevant_data)[names(all_relevant_data) == 'Q1r1'] <- 'Concern'
names(all_relevant_data)[names(all_relevant_data) == 'QD6'] <- 'Income'
names(all_relevant_data)[names(all_relevant_data) == 'QS2'] <- 'Gender'
names(all_relevant_data)[names(all_relevant_data) == 'QD5'] <- 'Education'
names(all_relevant_data)[names(all_relevant_data) == 'HIDAGE1'] <- 'Age'

# Filter out Don't know answers
all_relevant_data <- all_relevant_data %>% filter(!str_detect(`Concern`, "D"))
all_relevant_data <- all_relevant_data %>% filter(!str_detect(`Income`, "P"))
all_relevant_data <- all_relevant_data %>% filter(!str_detect(`Gender`, "P"))
all_relevant_data <- all_relevant_data %>% filter(!str_detect(`Gender`, "T"))
all_relevant_data <- all_relevant_data %>% filter(!str_detect(`Education`, "Prefer not to answer"))

# Transform Concern to Binary Variable
all_relevant_data$`Concern`[all_relevant_data$`Concern`=="Extremely concerned"] <- 1
all_relevant_data$`Concern`[all_relevant_data$`Concern`=="Very concerned"] <- 1

all_relevant_data$`Concern`[all_relevant_data$`Concern`=="Not very concerned"] <- 0
all_relevant_data$`Concern`[all_relevant_data$`Concern`=="Not at all concerned"] <- 0

all_relevant_data$`Age`[all_relevant_data$`Age` >= 18 & all_relevant_data$`Age` < 25] <- 1
all_relevant_data$`Age`[all_relevant_data$`Age` >= 25 & all_relevant_data$`Age` < 40] <- 2
all_relevant_data$`Age`[all_relevant_data$`Age` >= 40 & all_relevant_data$`Age` < 55] <- 3
all_relevant_data$`Age`[all_relevant_data$`Age` >= 55 & all_relevant_data$`Age` < 65] <- 4
all_relevant_data$`Age`[all_relevant_data$`Age` >= 65 & all_relevant_data$`Age` < 150] <- 5

all_relevant_data$`Income`[all_relevant_data$`Income`=="Under $40,000"] <- 1
all_relevant_data$`Income`[all_relevant_data$`Income`=="$100,001 to $150,000"] <- 2
all_relevant_data$`Income`[all_relevant_data$`Income`=="More than $150,000"] <- 3
all_relevant_data$`Income`[all_relevant_data$`Income`=="$40,001 to $60,000"] <- 4
all_relevant_data$`Income`[all_relevant_data$`Income`=="$80,001 to $100,000"] <- 5
all_relevant_data$`Income`[all_relevant_data$`Income`=="$60,001 to $80,000"] <- 6

all_relevant_data$`Gender`[all_relevant_data$`Gender` == "Man"] <- 1
all_relevant_data$`Gender`[all_relevant_data$`Gender` == "Woman"] <- 2

all_relevant_data$`Education`[all_relevant_data$`Education`=="High school or less"] <- 1
all_relevant_data$`Education`[all_relevant_data$`Education`=="Some community college, vocational, trade school"] <- 2
all_relevant_data$`Education`[all_relevant_data$`Education`=="Some university"] <- 2
all_relevant_data$`Education`[all_relevant_data$`Education`=="Completed community college, vocational, trade school"] <- 2
all_relevant_data$`Education`[all_relevant_data$`Education`=="Completed undergraduate degree"] <- 2
all_relevant_data$`Education`[all_relevant_data$`Education`=="Post graduate/professional school"] <- 3

all_relevant_data$Income <- as.numeric(as.character(all_relevant_data$Income))
all_relevant_data$Gender <- as.numeric(as.character(all_relevant_data$Gender))
all_relevant_data$Education <- as.numeric(as.character(all_relevant_data$Education))
all_relevant_data$Concern <- as.numeric(as.character(all_relevant_data$Concern))

logis_reg <- glm(Concern ~ Age + Gender + Education + Income,
                         data = all_relevant_data, family = "binomial")

```{r figmodel, echo = FALSE, fig.cap='Model Summary',fig.align="center", out.height = "250px", out.width = "250px"}
# call modelsummary(logis_reg) for the same result
knitr::include_graphics("./model_result.png")
```

# References
